---
title: "Graphs for Amita's paper - with data from U Laval"
author: "Karen Cristine Goncalves dos Santos"
date: "`r format(Sys.time(), '%b %d %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      root.dir = "E:/amita")
repos <- c("https://cran.rstudio.com/", "https://bioconductor.org/packages/3.11/bioc", 
					 "https://bioconductor.org/packages/3.11/data/annotation",
					 "https://bioconductor.org/packages/3.11/data/experiment")

for (i in c("magrittr", "reshape2", "tidyverse", "ggforce",
						"ensembldb", "ggdendro", "EnsDb.Hsapiens.v86", 
						"heatmap3", "lattice", "reshape2", "svglite")) {
	if (!i %in% rownames(installed.packages())) {
		install.packages(i, repos = repos)
	}
	library(i, character.only = T)
}

```
```{r echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
quants <- data.frame(Test = c(
  "IFN vs\nControl", "HIV1 vs\nControl", "HIV1+IFN vs\nControl", "HIV1+IFN\nvs IFN", "HIV1+IFN\nvs HIV-1"),
  Identified = rep(5223, 5),
  Quantifiable = c(4562, 4573, 4629, 4664, 4611),
  Quantified = c(4321, 4325, 4366, 4393, 4356)
  )

melt_quants <- melt(quants, id = "Test")
```

```{r include=FALSE}
org.adj.pVal <- read.delim("qval_zscore_stat.txt", header = T, stringsAsFactors = F)

qval <- org.adj.pVal[, c("Accession", grep("qvalue", colnames(org.adj.pVal), value = T))]
colnames(qval) <-  gsub("_qvalue", "", colnames(qval))

zscore_ratio <- org.adj.pVal[, c("Accession", grep("ZscRatio", colnames(org.adj.pVal), value = T))]
colnames(zscore_ratio) <- gsub("_ZscRatio", "", colnames(zscore_ratio))

status <- org.adj.pVal[, c("Accession", grep("stat", colnames(org.adj.pVal), value = T))]
colnames(status) <- gsub("_statVariant", "", colnames(status))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
melt_qval <- melt(qval, id = "Accession") %>% cbind(Variable = "qValue")
melt_zscore <- melt(zscore_ratio, id = "Accession") %>% cbind(Variable = "zscore_ratio")
melt_status <- melt(status, id = "Accession") %>% cbind(Variable = "status")
colnames(melt_qval)[2] <- colnames(melt_zscore)[2] <- colnames(melt_status)[2] <- "Test"

for_plots <- data.frame(melt_qval, melt_zscore[, "value"], melt_status[, "value"])
colnames(for_plots) <- c("Accession", "Test", "qValue", "variable", "Zscore_ratio", "Status")
for_plots$variable = NULL

for_plots$Test <- as.character(for_plots$Test) %>%
  gsub(pattern = "IFN_Virus", replacement = "HIV1+IFN") %>%
  gsub(pattern = "Virus", replacement = "HIV1") %>%
  gsub(pattern = "\\.", replacement = "\nvs ") %>%
  factor(levels = c("IFN\nvs Control", "HIV1\nvs Control", "HIV1+IFN\nvs Control", 
                    "HIV1+IFN\nvs IFN", "HIV1+IFN\nvs HIV1"))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
Total <- Down_regulated <- Up_regulated <- c()
summary_plot <- data.frame(matrix(nrow = 0, ncol = 3))
dap <- list()

for (i in levels(for_plots$Test)) {
  a <- gsub(replacement = "IFN_Virus", pattern = "HIV1\\+IFN", x = i) %>%
    gsub(replacement = "Virus", pattern = "HIV1") %>%
    gsub(replacement = "\\.", pattern = "\nvs ") 
  
  zsc <- paste0(a, "_ZscRatio"); qv <- paste0(a, "_qvalue"); statv <- paste0(a, "_statVariant")
  
  dap[[i]] <- org.adj.pVal[ abs(org.adj.pVal[, zsc]) > 1.96 & org.adj.pVal[, statv] == "VARIANT", "Accession" ]
  
  Total <- length(dap[[i]])
  
  Up_regulated <- length(zscore_ratio[ zscore_ratio$Accession %in% 
                                         dap[[i]] & zscore_ratio[, a] < -1.96, a])
  Down_regulated <- length(zscore_ratio[ zscore_ratio$Accession %in% 
                                           dap[[i]] & zscore_ratio[, a]  > 1.96, a])
  
  summary_plot <- rbind(summary_plot, c(Total, Up_regulated, Down_regulated)) %>% as.data.frame()
  colnames(summary_plot) <- c("Total", "Up-regulated", "Down-regulated")
  
  write.table(org.adj.pVal[abs(org.adj.pVal[, zsc]) > 1.96 & 
                             org.adj.pVal[, statv] == "VARIANT", 
                           c("Accession", zsc, qv, statv)], 
              paste0("Deregulated_proteins_", gsub("[\n ]", "_", i), ".txt"), 
              sep = "\t", quote = F, eol = "\n", row.names = F, col.names = T)

}
rownames(summary_plot) <- summary_plot$Test <- levels(for_plots$Test)
```


```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
normalized_abundances <- read.delim("normalized_abundances.txt", header = T, stringsAsFactors = F)
normalized_abundances$gene_name <- gsub(".*GN=(\\w+).*", "\\1", normalized_abundances$Description)
norm_abundances <- data.frame(normalized_abundances[,3:14] %>% log2(), 
                              row.names = normalized_abundances$Accession)

plot_boxplot <- melt(norm_abundances)
plot_boxplot$Test <- plot_boxplot$variable <- as.character(plot_boxplot$variable)

for (i in 1:nrow(plot_boxplot)) {
  if (grepl("CTRL", plot_boxplot$Test[i])) {
    plot_boxplot$Test[i] <- "Control"
    plot_boxplot$variable[i] <- gsub("CTRL", "CTRL_", plot_boxplot$variable[i])
  } else if (grepl("^IFN\\d$", plot_boxplot$Test[i])) {
    plot_boxplot$Test[i] <- "IFN"
    plot_boxplot$variable[i] <- gsub("IFN", "IFN_", plot_boxplot$variable[i])
  } else if (grepl("^VIRUS\\d$", plot_boxplot$Test[i])) {
    plot_boxplot$Test[i] <- "HIV1"
    plot_boxplot$variable[i] <- gsub("VIRUS", "HIV1_", plot_boxplot$variable[i])
  } else {
    plot_boxplot$Test[i] <- "HIV1+IFN"
    plot_boxplot$variable[i] <- gsub("IFN_VIRUS", "HIV1+IFN_", plot_boxplot$variable[i])
  }
}
plot_boxplot$Normalization <- "After Normalization"

```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
original_abundances <- read.delim("original_abundances.txt", header = T, stringsAsFactors = F)
original_abundances <- original_abundances[original_abundances$Accession %in% normalized_abundances$Accession, ]
colnames(original_abundances)[-1] <- 
  paste0(gsub("Abundance\\.\\.F\\w+\\.\\.Sample\\.\\.", "", colnames(original_abundances)[-1]), 1:3)

org_abundances <- data.frame(original_abundances[,2:13] %>% log2(), 
                              row.names = original_abundances$Accession)

plot_boxplot_org <- melt(org_abundances)
plot_boxplot_org$Test <- plot_boxplot_org$variable <- as.character(plot_boxplot_org$variable)

for (i in 1:nrow(plot_boxplot_org)) {
  if (grepl("CTRL", plot_boxplot_org$Test[i])) {
    plot_boxplot_org$Test[i] <- "Control"
    plot_boxplot_org$variable[i] <- gsub("CTRL", "CTRL_", plot_boxplot_org$variable[i])
  } else if (grepl("^IFN\\d$", plot_boxplot_org$Test[i])) {
    plot_boxplot_org$Test[i] <- "IFN"
    plot_boxplot_org$variable[i] <- gsub("IFN", "IFN_", plot_boxplot_org$variable[i])
  } else if (grepl("^VIRUS\\d$", plot_boxplot_org$Test[i])) {
    plot_boxplot_org$Test[i] <- "HIV1"
    plot_boxplot_org$variable[i] <- gsub("VIRUS", "HIV1_", plot_boxplot_org$variable[i])
  } else {
    plot_boxplot_org$Test[i] <- "HIV1+IFN"
    plot_boxplot_org$variable[i] <- gsub("IFN\\.VIRUS", "HIV1+IFN_", plot_boxplot_org$variable[i])
  }
}
plot_boxplot_org$Normalization <- "Before Normalization"

```


# 2) BoxPlot before and after normalization.
```{r echo=FALSE, fig.height=6, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
plot_boxplot_org$variable <- factor(plot_boxplot_org$variable, 
                                levels = c(paste0("CTRL_", 1:3), paste0("IFN_", 1:3),
                                           paste0("HIV1+IFN_", 1:3), paste0("HIV1_", 1:3)))
plot_boxplot_org$Test <- factor(plot_boxplot_org$Test, levels = c("Control", "IFN", "HIV1+IFN", "HIV1"))

joined_plots <- rbind(plot_boxplot, plot_boxplot_org)
joined_plots$Normalization <- factor(joined_plots$Normalization, 
                                     levels = c("Before Normalization", "After Normalization"))

ggplot(joined_plots, aes(variable, value, fill = Test)) + 
  geom_boxplot() + scale_fill_manual(values = c("#db8823", "#106909", "#8f2424", "#1082b3")) +
  labs(x = "", y = expression(log[2](Intensity))) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, color = "black", hjust = 1, vjust = 0.5),
        legend.position = "none") + 
  facet_wrap(~Normalization, nrow = 1)
ggsave("Comparison_before_after_norm_horizontal.tiff",
       device = "tiff", width = 10, height = 6, units = "in", dpi = 300)

ggsave("Comparison_before_after_norm_horizontal.eps",
       device = "eps", width = 10, height = 6, units = "in", dpi = 300)
```

```{r echo=FALSE, fig.height=10, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}

ggplot(joined_plots, aes(variable, value, fill = Test)) + 
  geom_boxplot() + scale_fill_manual(values = c("#db8823", "#106909", "#8f2424", "#1082b3")) +
  labs(x = "", y = expression(log[2](Intensity))) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, color = "black", hjust = 1, vjust = 0.5),
        legend.position = "none") + 
  facet_wrap(~Normalization, nrow = 2)

ggsave("Comparison_before_after_norm_vertical.tiff",
       device = "tiff", width = 6, height = 10, units = "in", dpi = 300)
ggsave("Comparison_before_after_norm_vertical.eps",
       device = "eps", width = 6, height = 10, units = "in", dpi = 300)

```

# 3) PCA analysis
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
pca_data <- data.frame(Replicate = plot_boxplot$variable %>% unique(),
                       Sample = gsub("(\\w+)\\d", "\\1", plot_boxplot$variable %>% unique()),
                       PC1 = c(51.2024, 36.6148, 34.5482, 49.1687, 30.2183, 27.3352, 
                               -0.6656, 9.3559, -59.22, -45.736, -71.3252, -52.4968),
                       PC2 = c(-10.2013, 24.966, 32.2231, -47.1303, -37.953, -8.705, 
                               14.1425, 59.481, 16.1858, -15.1009, -16.0312, -11.8766))

ggplot(pca_data, aes(PC1, PC2, color = Sample, label = Replicate)) + 
  labs(x = "PC1: 35.4% expl. variation", y = "PC2: 15% expl. variation") + 
  scale_color_manual(values = c("#1082b3", "#db8823", "grey3", "#106909")) + 
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) + 
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5) +
  geom_text(aes(PC1 + 2, PC2 - 4), size = 3) + geom_point() +
  #scale_x_continuous(breaks = seq(-75, 50, 12.5)) + 
  theme_bw() + theme(legend.position = "none",
                     axis.text = element_text(color = "black"),
                     aspect.ratio = 0.5)
ggsave("PCA.tiff", device = "tiff", width = 7, height = 4, units = "in", dpi = 300)
ggsave("PCA.eps", device = "eps", width = 7, height = 4, units = "in", dpi = 300)

```

# 4) Number of quantifications for each sample
```{r echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(melt_quants, aes(Test, value, fill = variable, color = variable)) + 
  geom_col(position = "dodge2") + scale_fill_discrete(name = "") + 
  scale_color_discrete(name = "") +
  geom_text(aes(label = value, y = value - 290), 
            angle = 90, position = position_dodge(0.9), color = "black") +
  labs(x = "Test", y = "Number of proteins") +
  theme_bw() + theme(axis.text = element_text(color = "black"),
                     legend.position = "bottom",)

ggsave("Quant_prot.tiff", device = "tiff", width = 6, height = 6, units = "in", dpi = 300)
ggsave("Quant_prot.eps", device = "eps", width = 6, height = 6, units = "in", dpi = 300)
```

# 5) Number of variant proteins for each comparison
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot_summary <- reshape2::melt(summary_plot, id = "Test")

plot_summary$Test <- factor(plot_summary$Test, levels = levels(for_plots$Test))

ggplot(plot_summary, aes(Test, value, fill = variable, color = variable)) +
  geom_col(position = "dodge2") + 
  geom_text(aes( label = value, y = value + 10), position = position_dodge(0.9), color = "black") + 
  scale_color_manual(name = "", values = c("darkgrey", "red", "navyblue"))  +
  scale_fill_manual(name = "", values = c("darkgrey", "red", "navyblue")) + 
  labs(y = "Number of variant proteins", x = "Comparison") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, color = "black", hjust = 1, vjust = 0.5),
        axis.text.y = element_text(color = "black"),
        legend.position = "bottom")
ggsave("Variant_prot.tiff",device = "tiff", width = 8, height = 6, units = "in", dpi = 300)
ggsave("Variant_prot.eps",device = "eps", width = 8, height = 6, units = "in", dpi = 300)

```


```{r echo=FALSE, fig.height=10, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ComplexHeatmap)
mean <- apply(normalized_abundances[3:14], 1, mean)
stdev <- apply(normalized_abundances[3:14], 1, sd)

heatmap_data <- data.frame(cbind((normalized_abundances[3:14] - mean) / stdev, 
                                 Accession = normalized_abundances$Accession))
colnames(heatmap_data)[1:12] <- plot_boxplot$variable %>% unique()

dend_samples <- as.matrix(normalized_abundances[3:14]) %>% t() %>% dist() %>% hclust() 
order_samples <- dend_samples$order

dend_prots <- as.matrix(normalized_abundances[3:14]) %>% dist() %>% hclust() 
order_prots <- dend_prots$order

```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
Tests <- for_plots
Tests$Status <-  as.character(Tests$Status)
for (prot in 1:nrow(Tests)) {
    
    if (Tests$Status[prot] == "VARIANT") {
      
      if (Tests$Zscore_ratio[prot] > 1.96 ) {
        Tests$Status[prot] <- "Down-regulated"
      } else if (Tests$Zscore_ratio[prot] < -1.96) {
        Tests$Status[prot] <- "Up-regulated"
      } else {
        Tests$Status[prot] <- FALSE
      }
      
    } else if (!is.na(Tests$qValue[prot]) & Tests$qValue[prot] < 0.05) {
      
      if (Tests$Zscore_ratio[prot] > 1.96 ) {
        Tests$Status[prot] <- "Down-regulated"
      } else if (Tests$Zscore_ratio[prot] < -1.96) {
        Tests$Status[prot] <- "Up-regulated"
      } else {
        Tests$Status[prot] <- FALSE
      }
      
    } else {
      Tests$Status[prot] <- FALSE
    }
}
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
edb <- EnsDb.Hsapiens.v86::EnsDb.Hsapiens.v86

txs <- transcripts(edb, columns = c("gene_name", "uniprot_id"))@elementMetadata@listData %>%
  as.data.frame() %>% dplyr::select(uniprot_id, gene_name) 

txs <- dplyr::filter(txs, !is.na(uniprot_id) & !is.na(gene_name))

gene_names <- data.frame(0, 0)
for (i in org.adj.pVal$Accession) { 
  if (i %in% gene_names$X0) {next}
  gene_names <- rbind(gene_names, c(X0 = i, X0.1 = as.character(txs$gene_name[txs$uniprot_id == i][1])))
}

gene_names <- gene_names[-1,]
colnames(gene_names) <- colnames(txs)
```

# 6) Volcano plots

```{r echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
vplots_data <- list()

for (i in unique(Tests$Test)) {
  a <- gsub("\n", " ", i)
  if (i == "HIV1\nvs Control") {
    data <- dplyr::filter(Tests, Test == i) %>% cbind(gene_name = gene_names$gene_name) 
    
    vplots_data[[a]] <- 
      ggplot(data, aes(-1 * Zscore_ratio, -log10(qValue), color = Status, label = gene_name)) + 
      geom_point(size = 1) + 
      geom_text(data =  dplyr::filter(data, Status != FALSE), 
                nudge_y = c(0.3, -0.3), size = 2) +
      geom_point(aes(x = -1 * (data$Zscore_ratio[data$Accession == "P12270"]), 
                     y = -log10(data$qValue[data$Accession == "P12270"])),
                 size = 3, color = "black") +
      scale_color_manual(values = c("blue", "grey", "red")) + 
      labs(x = expression(log[2]("Z-Score Ratio")), 
           y = expression(-log[10](qValue)), 
           title = a) + 
      theme_bw() + theme(legend.position = "none") 
  } else {
    data <- dplyr::filter(Tests, Test == i) %>% cbind(gene_name = gene_names$gene_name) 
    
    vplots_data[[a]] <- 
      ggplot(data, aes(-1 * Zscore_ratio, -log10(qValue), color = Status, label = gene_name)) + 
      geom_point(size = 1) + 
      geom_text(data =  dplyr::filter(data, Status != FALSE), 
                nudge_y = c(0.3, -0.3), size = 2.5) +
      scale_color_manual(values = c("blue", "grey", "red")) + 
      labs(x = expression(log[2]("Z-Score Ratio")), 
           y = expression(-log[10](qValue)), 
           title = a) + 
      theme_bw() + theme(legend.position = "none") 
  }
  
  print( vplots_data[[a]])
  ggsave(plot = vplots_data[[a]], paste0("Volcano_plot", a, ".tiff"), 
         device = "tiff", dpi = 300)
  ggsave(plot = vplots_data[[a]], paste0("Volcano_plot", a, ".eps"), 
         device = "eps", dpi = 300)
  dev.off()

}
```

# 7) Heatmaps

## Single heatmap with all proteins (deregulated or not)

```{r echo=FALSE, fig.height=12, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
heat <- Heatmap(heatmap_data[, -13], name = "Zscore ratio", column_names_side = "top",
        cluster_columns = F, show_row_dend = F, cluster_rows = T,
        show_column_dend = F, column_dend_side = "top",
        row_labels = rep("", nrow(heatmap_data)),
        row_title = "Dendrogram calculated with Zscore")

draw(heat)
postscript(file = "Heatmap_all_prots.eps")
heat
dev.off()

tiff(filename = "Heatmap_all_prots.tiff", width = 7, height = 12, units = "in", res = 300)
heat 
dev.off()
```


## Single heatmap with only deregulated proteins 

```{r echo=FALSE, fig.height=12, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
deregulated_prots <- union(dap[[1]], union(dap[[2]], union(dap[[3]], union(dap[[4]], dap[[5]]))))
heat <- Heatmap(heatmap_data[heatmap_data$Accession %in% deregulated_prots, -13], 
                name = "Zscore ratio", column_names_side = "top",
                cluster_columns = F, show_row_dend = F, cluster_rows = T,
                show_column_dend = F, column_dend_side = "top",
                row_labels = rep("", length(deregulated_prots)),
                row_title = "Dendrogram calculated with Zscore")
draw(heat)

tiff(filename = "Heatmap_deregulated_prots.tiff", width = 7, height = 12, units = "in", res = 300)
heat
dev.off()

postscript(file = "Heatmap_deregulated_prots.eps")
heat
dev.off()

```

```{r echo=FALSE, fig.height=12, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
heatmap_dt <- data.frame(heatmap_data[, -13], 
  gene_names, row.names = heatmap_data$Accession)
heatmap_dt <- heatmap_dt[, !colnames(heatmap_dt) %in% c("Accession", "uniprot_id")]
colnames(heatmap_dt) <- c(colnames(heatmap_data)[-13],  "gene_name")
deregs <- samples <- data <- dereg_heatmaps <- name <- fontface <- list()
for (i in levels(Tests$Test)) {
  if (i %in% c("IFN\nvs Control", "HIV1\nvs Control")) {
    deregs[[i]] <- Tests$Accession[Tests$Test == i & Tests$Status %in% c("Up-regulated", "Down-regulated")]
    name[[i]] <- heatmap_dt$gene_name[rownames(heatmap_dt) %in% deregs[[i]]]
    
    for (prot in 1:length(name[[i]])) {
      name[[i]][prot] <- ifelse(is.na(name[[i]][prot]),
                                deregs[[i]][prot],
                                name[[i]][prot])
      
    }
    
    samps <- paste0(c("HIV1_", "CTRL_", "IFN_"), 
                    c(rep(1, 3), rep(2, 3), rep(3, 3))) %>% sort()
    samples[[i]] <- samps[1:3]
    if (grepl("HIV", i)) {
      samples[[i]] <- c(samples[[i]], samps[4:6])
    } else {
      samples[[i]] <- c(samples[[i]], samps[7:9])
    }
      
    data[[i]] <- heatmap_dt[deregs[[i]], samples[[i]]]
    dereg_heatmaps[[i]] <- Heatmap(data[[i]], row_names_gp = gpar(fontsize = 5),
                                   column_title = gsub("\n", " ", i), 
                                   column_names_gp = gpar(fontsize = 9),
                                   cluster_columns = F,
                                   column_names_side = "top", row_names_side = "left",
                                   row_labels = name[[i]], name = "log2(Intensity)"
                                   )
  }
}
```

## Heatmap of deregulated proteins in test *IFN vs Control*

```{r echo=FALSE, fig.height=12, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}

dereg_heatmaps[[1]]

tiff(filename = "Heatmap_deregulated_IFNvsControl.tiff", width = 7, height = 12, units = "in", res = 300)
dereg_heatmaps[[1]]
dev.off()

postscript(file = "Heatmap_deregulated_IFNvsControl.eps")
dereg_heatmaps[[1]]
dev.off()

```

## Heatmap of deregulated proteins in test *HIV-1 vs Control*

```{r echo=FALSE, fig.height=12, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
dereg_heatmaps[[2]]
tiff(filename = "Heatmap_deregulated_HIV-1vsControl.tiff", width = 7, height = 12, units = "in", res = 300)
dereg_heatmaps[[2]]
dev.off()
postscript(file = "Heatmap_deregulated_HIV-1vsControl.eps")
dereg_heatmaps[[2]]
dev.off()

```

# 10) HIV-1 peptides in the nuclear membrane
```{r echo=FALSE, fig.height=4, fig.width=4, message=FALSE, warning=FALSE, paged.print=FALSE}
HIV_prots <- data.frame(
  "Protein_region" = c(rep("P17", 6), rep("P24 N terminus", 4), "P24 (N terminus, Linker, C terminus)", rep("P24  C terminus", 2), "P6"),
  "Peptides" = c("(R)ASVLSGGELDK(W)", "(R)ASVLSGGELDKWEK(I)", "(R)FAVNPGLLETSEGcR(Q)", "(R)QILGQLQPSLQTGSEELR(S)", "(R)SLYNTIAVLYcVHQR(I)", "(K)EALDKIEEEQNK(S)", "(K)ETINEEAAEWDR(L)", "(R)LHPVHAGPIAPGQMR(E)", "(R)GSDIAGTTSTLQEQIGWMTHNPPIPVGEIYK(R)", "(R)WIILGLNK(I)", "(R)MYSPTSILDIR(Q)", "(K)NWMTETLLVQNANPDcK(T)", "(K)ALGPGATLEEMMTAcQGVGGPGHK(A)", "(R)QEPIDKELYPLASLR(S)"),
  "HIV1_1" = c(0, 2, 1, 1, 1, 2, 1, 0, 0, 1, 0, 0, 0, 0), 
  "HIV1_2" = c(0, 2, 1, 1, 1, 1, 2, 0, 0, 1, 1, 0, 0, 0),
  "HIV1_3" = c(0, 1, 1, 1, 0, 1, 0, 2, 0, 0, 0, 0, 0, 0),
  "HIV1_IFNb_1" = c(1, 2, 0, 1, 1, 2, 1, 3, 1, 1, 2, 1, 1, 1),
  "HIV1_IFNb_2" = c(0, 2, 1, 1, 0, 2, 0, 4, 1, 1, 1, 0, 1, 0),
  "HIV1_IFNb_3" = c(0, 2, 1, 1, 1, 2, 1, 3, 1, 0, 1, 0, 1, 0), stringsAsFactors = F)

t_test <- list()
for (i in unique(HIV_prots$Protein_region)) {
  t_test[[i]] <- data.frame(HIV = c(HIV_prots[HIV_prots$Protein_region == i,3], 
                                    HIV_prots[HIV_prots$Protein_region == i,4],
                                    HIV_prots[HIV_prots$Protein_region == i,5]), 
                            HIV_IFN = c(HIV_prots[HIV_prots$Protein_region == i,6],
                                        HIV_prots[HIV_prots$Protein_region == i,7], 
                                        HIV_prots[HIV_prots$Protein_region == i,8])) %>% t.test()
}

t_test_all <- list()
for (i in colnames(HIV_prots[, 3:8])) {
  if (grepl("IFNb", i)) {
    t_test_all[["HIV1_IFNb"]] <- c(t_test_all[["HIV1_IFNb"]], sum(HIV_prots[, i]))
  } else {
    t_test_all[["HIV1"]] <- c(t_test_all[["HIV1"]], sum(HIV_prots[, i]))
  }
}

t_test_all <- data.frame(HIV_IFNb =  t_test_all[["HIV1_IFNb"]], HIV1 = t_test_all[["HIV1"]])
t.test(t_test_all)
HIV_prots$HIV1 <- rowSums(HIV_prots[, 3:5])
HIV_prots$HIV1_IFNb  <- rowSums(HIV_prots[, 6:8])

ggplot(data = data.frame(variable = c("HIV1", "HIV1 + IFNb"), value = apply(HIV_prots[, c("HIV1", "HIV1_IFNb")], 2, sum)), 
       aes(variable, value, fill = variable)) + 
  geom_text(aes(variable, value + 2, label = value)) + 
  geom_col() + labs(x = "", y = "Total HIV-1 unique spectra (3x exp.)") + 
  scale_fill_manual(values = c("#106909", "#8f2424")) + 
  scale_y_continuous(limits = c(0, 50)) + theme_bw() + 
  theme(legend.position = "None", 
        aspect.ratio = 1.5, 
        axis.title = element_text(size = 12, color = "black"), 
        axis.text = element_text(size = 12, color = "black"),
        plot.title = element_text(hjust = 0.5))

ggsave(paste0("HIV_pep_membrane", ".tiff"), device = "tiff", width = 4, height = 4, units = "in", dpi = 300)
ggsave(paste0("HIV_pep_membrane", ".eps"), device = "eps", width = 4, height = 4, units = "in", dpi = 300)

```

```{r}
interaction_hiv <- read.delim("Karen vs ULaval and inetraction with HIV-1 protein.txt", 
                              header = T, row.names = 2, stringsAsFactors = F)
interactions <- interaction_hiv[interaction_hiv$X == "", -c(1, 3, 4)]

for (j in 2:ncol(interactions)) {
  for (i in 1:nrow(interactions)) {
    interactions[i, j] <- ifelse(is.na(interactions[i, j]), 0, 
                                 ifelse(interactions[i, j] == "", 0, 1))
  }
  interactions[, j] <- as.numeric(interactions[, j])
}

melt_interactions <- melt(interactions, id = "U.laval.Accession")
melt_interactions <- melt_interactions[melt_interactions$value == 1,-3]
melt_interactions$variable <- as.character(melt_interactions$variable) %>%
  gsub(pattern = "Retropepsin", replacement = "Protease")
names(melt_interactions) <- c("Human", "HIV-1")

nodes <- data.frame(Protein_ID = c(unique(melt_interactions$`HIV-1`), 
                                   unique(melt_interactions$Human)),
                    Origin = c(rep("HIV-1", unique(melt_interactions$`HIV-1`) %>% length),
                               rep("Human", unique(melt_interactions$Human) %>% length)),
                    Protein_name = c(unique(melt_interactions$`HIV-1`), 
                                     gene_names[gene_names$uniprot_id %in% melt_interactions$Human, 2]))
hiv <- c()
for (i in 1:nrow(nodes)) {
  hiv <- c(hiv, ifelse(nodes$Protein_ID[i] %in% 
                         vplots_data$`HIV1 vs Control`$data$Accession, 
                       vplots_data$`HIV1 vs Control`$data$Zscore_ratio[
                         vplots_data$`HIV1 vs Control`$data$Accession == nodes$Protein_ID[i]], 0))
}

nodes$HIV1.vs.Control <- ifelse(hiv == 0, 0, ifelse(hiv > 0, "Up-regulated", "Down-regulated"))
nodes$width <- str_count(nodes$Protein_name, "")*8 + 20

write.table(x = nodes, file = "nodes_interactions.txt", 
            quote = F, sep = "\t", eol = "\n", row.names = F, col.names = T)
write.table(x = melt_interactions, file = "edges_interactions.txt", 
            quote = F, sep = "\t", eol = "\n", row.names = F, col.names = T)
```

